---
title: |
       | Supplementary Materials for "Legislature Size and Welfare: Evidence from Brazil"
author:
  - Umberto Mignozzetti[^umberto]
  - Gabriel Cepaluni[^gabriel]
  - Danilo Freire[^danilo]
date: \today
fontfamily: libertine
fontawesome: yes
fontsize: 11pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
always_allow_html: true
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: true
    keep_tex: no
    template: template.latex
---

[^umberto]: Visiting Assistant Professor, Quantitative Theory and Methods Department, Emory University, <umberto.mignozzetti@emory.edu>, <http://umbertomig.com>. Corresponding author.

[^danilo]: Senior Lecturer, School of Social and Political Sciences, University of Lincoln, <danilofreire@gmail.com>, <https://danilofreire.github.io>.

[^gabriel]: Associate Professor, Department of International Relations, São Paulo State University, <gabi.cepal@gmail.com>.

\appendix
\addtocounter{section}{2}

```{r, include=FALSE}
# Knitr options
knitr::opts_chunk$set(fig.pos = "H") # holds figure position
knitr::opts_chunk$set(echo = TRUE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

```{r, include=FALSE}
## Starting
#XXXX

# Required packages
#XXXX

# Load datasets
#XXXX
```

\newpage

# A Model of Legislature Size and Service Provision

## Full model

Consider an strategic interaction comprised of a mayor ($M$), and $N \geq 3$ of city councilors, indexed in $i \in \{1, 2, \cdots, N \}$. The mayor is head of the local Executive, and has the prerogative of proposing a vector of policies, that is voted by the city councilors. The policy proposal, if accepted, is implemented.

Every city councilor has a party affiliation, and the party label may or may not match the mayor's affiliation. If the politician shares the same party as the mayor, we say that it belongs to the government coalition ($G \subset \{1, 2, \cdots, N \}$). Otherwise, we assume that the politician belongs to the opposition $O \not\subset G$. The party affiliation is important because it provides a political benefit or cost for the councilor. If the mayor implements the policy, then a councilor aligned with the mayor receives a policy benefit $p > 0$. If the politician belongs to the opposition, then the mayoral policies impose her a cost of $-p$. For simplicity, we assume that these costs are symmetric.

However, if the mayor's policy proposal lose majoritarian support from the council, then we assume that a reversal policy takes place, where the resources diminish by a factor of $\delta \in (0,1)$, and a random councilor is recognized to make a proposal.\footnote{In a model similar to [Baron Ferejohn / Cliff XXXX].} The councilor proposes the new policy vector, and the council votes. If the policy is rejected, then another councilor is recognized to make the offer, and this goes on until one offer is finally accepted by the majority of the legislators. For simplicity, we assume that the councilor's offers generate no political benefits or costs, so that if a councilor proposes the policy vector, then $p=0$. This may sound restrictive, but we will show that councilors propose zero public goods in equilibrium, as they have only to please their particularistic constituencies to get reelected.

To convince the councilors, the politician making the proposal has three vectors of policies at her disposal. First, the proposer may give targeted goods for the councilor $i$ that she is trying to earn support, $t_i \geq 0$. These targeted goods improve the reelection rates of the councilor. Second, the proposer may also give private goods $x_i \geq 0$ to a councilor $i$. Private goods do not affect the reelection rates of the politicians, but provide instant benefits for the receiver. Finally, the proposer can provide public goods $g \geq 0$. Public goods help the mayor to get reelected, and also affect the policy benefits received by the politicians, depending on the party affiliation. 

For the mayor, the expected utility function is a sum of the utility from rents and benefits from reelection. Both the utilities from rents ($u(r)$) and the probability of getting reelected $\pi(g)$ are concave functions, meaning that more rents or public goods increase the utility at a decreasing rate. The probability of getting reelected is also multiplied by the benefit from holding office $B_M > 0$. This term captures the tangible and intangible gains that the mayor perceives from being reelected into the position. The utility is multiplied by an indicator $\mathbb{I}$ that captures whether the mayor's proposal was accepted by the city council.\footnote{The indicator function is defined as
\[
I(condition) \ = \ \twopdo{1}{\text{Condition is satisfied.}}{0}.
\]
} The expected utility for the mayor is:

\[
\mathbb{E}U(r, g, \bf{x}, \bf{t}) \ = \ \mathbb{I}(\text{Approval})(u(r) + B_M\pi(g))
\]

And the mayor chooses the optimal policy subject to a budget constraint. The municipality has $R$ resources, and we assume that the municipality cannot run debts. This budget balance constraint requires the policy vector $(r, g, \bf{x}, \bf{t})$ to satisfy the following rule:

\[
r + g + \sum_{i=1}^{N}t_i + \sum_{i=1}^{N} x_i \ \leq \ \delta^kR
\]

And the mayor's proposal happens when $k=0$. This makes mayoral budget inequality to satisfy: $r + g + \sum_{i=1}^{N}t_i + \sum_{i=1}^{N} x_i \ \leq \ R$. Otherwise, in the reversal proposals, the resources diminish by a $\delta^k$ rate, with $k$ being the number of unsuccessful offers by the councilors. 

We can decompose the councilor's utility function into three major components. First, the political gains or losses (depending on the councilor's party affiliation); second, the rents $x_i$ offered by the policy proposer to the councilor; and finally, the transfers for the constituents $t_i$, offered by the policy proposer to the councilor. As the constituent transfer $t_i$ affects chance of winning the election $\pi_C(t_i)$, we multiply this chance by the benefits accrued from remaining in office, $B > 0$. Again, as in the mayor's case, the utility is only realized when the city council accepts the policy proposed, so that we multiply the approval indicator $\mathbb{I}(.)$. The councilor's expected utility is equal to:

\[
\mathbb{E}U(r, g, \bf{x}, \bf{t}) \ = \  \mathbb{I}(\text{Approval})((-1)^{\mathbb{I}(i \in O)}p + x_i + B\pi_C(t_i))
\]

For simplicity, we assume that the councilor's probability of reelection satisfy a simple cutoff rule:

\[
\pi_C(t_i) \ = \ \twopdo{1}{t_i \geq \underline{t}}{0}
\]

This assumption simplifies the calculations, and it is reasonable as many models have the voter's decision as a cutoff strategy [@XXXX]. We also assume, for convenience, that $B > \underline{t}$. Otherwise, the councilor would prefer to pocket the targeted goods, and not being reelected.

The timeline of the game is as follows:

\begin{enumerate}
  \item The mayor learns how many government $|G|$ and opposition $|O|$ legislators were elected.
  \item The mayor proposes a policy vector $(r, g, \bf{x}, \bf{t})_M$.
  \item The city council votes the proposal.
  \begin{itemize}
    \item If the council accepts, then the policy is implemented, and the game ends. 
    \item Otherwise, the reversal policy is implemented.
  \end{itemize}
  \item (Reversal Policy) At the beginning of each iteration of the reversal policy, one councilor is randomly recognized to be the proposer. She proposes a vector $(\bf{x}, \bf{t})$, with $r =  g = 0$. If the proposal is accepted, it is implemented. If the proposal is rejected, a new round of reversal start, but with the municipal resources reduced by a factor $\delta$.
\end{enumerate}

This is a bargaining game, along the lines of @rubinsteinXXXX. For political science models similar to this model, see @baronfereXXXX and @cliffsXXXX. 

The solution for this game consists in finding a Stationary Sub-game Perfect Nash Equilibrium using backward induction. Although the game tree may extend infinitely, a Stationary Sub-game Perfect Equilibrium assumes that if a politician accepts an offer at time $t+1$, she should accept the offer at time $t$. This gives us an objective of finding the offer that will be accepted at any point in the game. Moreover, applying this reasoning we can find the optimal offer at $t=0$, which represents no delay in policy implementation. The mayor calibrates the offer for approval offering the benefits that councilors would receive, should they reject the offer and enter the reversal stage.

\begin{definition}
The solution for the game consists in a pair of strategic policy vectors for the mayor, and the city councilors. The mayor's policy vector is comprised of a public goods provision $g^*$ and XXXX
\end{definition}

Using backward induction, suppose that there were $k$ prior rejections, and the municipal resources shrank by $\delta^{k}R$. At this stage, the politicians are similar with regard to political affiliation, so that we drop the party label. A given councilor is recognized as the proposer with $1/N$ probability. The recognized councilor has to convince at least one-half of the council members to agree with her policy proposal. Therefore, a legislator $i$, that receives an offer from a proposer ($x_i, t_i$), at round $k$, will accept the offer if it is at least as good as waiting for the next round.

\[
x_i + B\pi_C(t_i) \ \geq \ \dfrac{1}{N}\left[\delta^{k+1}R - \dfrac{N}{2}(x_{-i} + t_{-i}) + B\pi_C(t_i)\right] + \dfrac{1}{2}(x_i + B\pi_C(t_i))
\]

At the $k$-th stage, the players are equivalent. Therefore, the offer $x_i, t_i$, and the offer for the other players, ($x_{-i} + t_{-i}$), are the same. Moreover, as $B > \underline{t}$, $t^* = t_i = t_{-i} = \underline{t}$, and the offer that would make legislator $i$ accept without further delay is:

\[
x + B \ \geq \ \dfrac{1}{N}\left(\delta^{k+1}R - \dfrac{N}{2}(x + \underline{t}) + B\right) + \dfrac{1}{2}(x + B)
\]

And after some algebra, the optimal private goods offer has to satisfy:

\[
x \geq \dfrac{1}{N}\left( \delta^{k+1}R + B \right) - \dfrac{1}{2}(B + \underline{t}) \equiv \underline{x}(k)
\]

For all $k \in \{1, 2, \cdots, \infty\}$.

\begin{lemma}
Suppose that the council rejected the mayor's offer. Then, at the $k$-th stage of the game, the proposer will offer $t^* = \underline{t}$ and $x^* = \underline{x}(k)$ for half of the members, keeping for her $t_i = \max\{0, \underline{t}\}$, and

\[
x_i \ = \ R\delta^k\left( 1 - \dfrac{\delta}{2} \right) + \dfrac{N}{4}(B-\underline{t}) + \dfrac{B}{2}
\]

Subject to the budget constraint:

\[
\dfrac{N}{2}\left(\underline{t}+\underline{x}(k)\right) + x_i  \ \leq \ \delta^kR
\]

The sequence of strategies derived here are an equilibrium for all $k \in \{1, 2, \cdots, \infty\}$.

\end{lemma}

\begin{proof}
The proof is direct, after collecting the terms from the previous discussion.

Note that after some iterations, the budget may shrink enough so that $t^* = \underline{t}$ is unfeasible. This means that there exists a $k^*$ such that, when $k > k^*$, $t^* = 0$, and the councilors divide the remaining budget as rents. This case is trivial to solve, making $\underline{x}(k) = \dfrac{1}{N}\left(\delta^{k+1}R\right)$ and $x_i = \delta^kR\left(1 - \dfrac{\delta}{2}\right)$.
\end{proof}

We now have the entire strategy profile for the city councilors, provided that they reject the initial mayor's proposal. Proceeding by backward induction, we solve for the mayor's proposal that wins the support of the majority of the council. In the mayor's proposal policy concerns come into play and the offer for a government-aligned councilor is different than the offer for an opposition councilor. For a government aligned councilor, assuming the councilor is the pivotal for the implementation, the accepted offer $(x_m^g, t_m^g)$ has to be better than rejecting the proposal, and waiting for the next round:

\[
p + x_m^g + B\pi_C(t_m^g) \ \geq \ \dfrac{1}{N}\left( \delta R -\dfrac{N}{2}(\underline{x}(1) + \underline{t}) + B\pi_C(t_i))  \right) + \left(1 - \dfrac{1}{N}\right)\left(\dfrac{1}{2}\right)(\underline{x}(1) + B\pi_C(t_i))
\]

Note that $\underline{x}(1)$ is the offer of private goods if the councilor rejects the mayors offer, so that $k = 1$. As the benefit of retention $B$ is higher than the cost of providing the targeted goods $\underline{t}$, the optimal provision of targeted goods satisfy the threshold for reelection: $t^* = t_m^g = t_i = \underline{t}$. Rewriting the equations leads us to:

\[
p + x_m^g + B \ \geq \ \dfrac{1}{N}\left( \delta R -\dfrac{N}{2}(\underline{x}(1) + \underline{t}) + B)  \right) + \left(1 - \dfrac{1}{N}\right)\left(\dfrac{1}{2}\right)(\underline{x}(1) + B)
\]

And after some algebra:

\[
x_m^g \ \geq \ -p +\dfrac{1}{2N+1}\left[2\delta R \left(1 + \dfrac{1-\delta}{2N}\right)-N(\underline{t} + B) + B\right] \ \equiv \ \underline{x}_m^g
\]

For the opposition councilor, the offer has to compensate the political loss of having a mayor that is not aligned with the councilor's ideology. The optimal offer satisfy:

\[
-p + x_o^g + B \ \geq \ \dfrac{1}{N}\left( \delta R -\dfrac{N}{2}(\underline{x}(1) + \underline{t}) + B)  \right) + \left(1 - \dfrac{1}{N}\right)\left(\dfrac{1}{2}\right)(\underline{x}(1) + B)
\]

And after some algebra:

\[
x_o^g \ \geq \ p + \dfrac{1}{2N+1}\left[2\delta R \left(1 + \dfrac{1-\delta}{2N}\right)-N(\underline{t} + B) + B\right] \ \equiv \ \underline{x}_o^g
\]

\begin{lemma}
The offer for a government-aligned councilor is always lower than the offer for an opposition councilor.
\end{lemma}

\begin{proof}
Note that the $\underline{x}_m^g = 2p + \underline{x}_o^g$. As the political disagreement is higher than zero, $\underline{x}_m^g = - 2p + \underline{x}_m^o < \underline{x}_m^g$. Moreover, note that $\underline{x}_m^g - \underline{x}_m^o = -2p$.
\end{proof}

This characterizes the offer that the mayor will make to win the support of the council. The next Definition helps to decompose the politician's offer into two components: rents and policy compensation.

\begin{definition}
Define the rents offered from the bargaining process as

\[
b(N) \ = \ \dfrac{1}{2N+1}\left[2\delta R \left(1 + \dfrac{1-\delta}{2N}\right)-N(\underline{t} + B) + B\right]
\]

When the legislature size is $N$, the optimal offers for an opposition member is:

\[
\underline{x}_m^o(N) = b(N) + p
\]

And the optimal offer for a government member is:

\[
\underline{x}_m^g(N) = b(N) - p
\]

\end{definition}

Now, let the number of government politicians be equal to $\eta_g$, and the ex-ante probability of electing a councilor aligned with the government equal to $\gamma \in (0,1)$. We define the cost to govern $C_G(N)$ as three parts equation: first, the costs of making the electoral transfers for at least half of the councilors; second, the rents paid in bargaining; and last, the political costs, that depend on the amount of politicians that are affiliated with the mayor's party. After simplifications, the equation describing the costs of govern is as follows:

\[
C_G(N) = \dfrac{N}{2}(\underline{t} + b(N) + p) - 2p\sum_{\eta_g = 0}^{N/2}\mathbb{P}(\eta_g)\eta_g
\]

And the probability that $\eta_g$ councilors belong to the mayor's coalition is equal to $\mathbb{P}(\eta_g) = \binom{N}{\eta_g}\gamma^{\eta_g}(1-\gamma)^{N-\eta_g}$, which is the binomial probability given that the ex-ante chance to elect an aligned councilor is equal to $\gamma$.

We need now to derive the optimal rents ($r$) and public goods provision ($g$) proposed by the mayor. The mayor benefits from public goods provision, as it increases the chances of her reelection. However, she prefers to contribute as minimum as possible for the public goods, and extract the remaining resources as political rents. Therefore, her objective is to maximize the expected utility subject to the budget constraint.

\[
\begin{aligned}
\max_{r, g} \quad & u(r) + B_M\pi(g)\\
\textrm{s.t.} \quad & r + g + C_G(N) \leq R
\end{aligned}
\]

In equilibrium the budget binds, and the optimal public goods provision maximizes:

\[
\max_{g} \{ u(R - g - C_G(N)) + B_M\pi(g) \}
\]

The first order condition for an optimal public goods provision makes the marginal benefits to increase rents, in this case, marginal costs of providing the public goods, equals to the marginal benefits of reelection:

\[
u'(R - g - C_G(N)) = B_M\pi'(g)
\]

This is sufficient for the equilibrium, as the second order condition reassures the concavity of the mayor's expected utility: $u''(R - g^* - C_G(N)) + B_M\pi''(g^*) < 0$.

\begin{proposition}
The provision of public goods increase with legislature size if the costs of govern decreases in larger legislatures.
\end{proposition}

\begin{proof}
In order to prove this result, we need to show that the mayor's expected utility satisfy the increasing differences in $g$ and $N$. This would mean that when increasing the size of the council, the optimal solution $g^*$ would also increase. However, as $N$ increases discretely, we cannot take a derivative on $N$ and use the implicit function theorem. We use Monotone Comparative Statics [@milgromshannonXXXX] to derive these results.

The mayor's expected utility satisfy the increasing differences in $g$ and $N$ when, for $g > g'$ and $N+1>N$, we have:

\[
\mathbb{E}\text{U}(g, N+1) - \mathbb{E}\text{U}(g', N+1) \geq \mathbb{E}\text{U}(g, N) - \mathbb{E}\text{U}(g', N)
\]

And after substituting, we have:

\[
u(R - g - C_G(N+1)) + B_M\pi(g) - u(R - g' - C_G(N+1)) - B_M\pi(g')
\\
\geq 
\\
u(R - g - C_G(N)) + B_M\pi(g) - u(R - g' - C_G(N)) - B_M\pi(g')
\]

Which is equal to:

\[
u(R - g - C_G(N+1)) - u(R - g' - C_G(N+1)) \geq 
u(R - g - C_G(N)) - u(R - g' - C_G(N))
\]

Multiplying both sides by $\dfrac{1}{g - g'}$ and taking the limit when $g \rightarrow g'$, we have:

\[
u'(R - g - C_G(N+1)) \geq 
u'(R - g - C_G(N))
\]

This is true when the cost of govern with a council size of $N+1$ is lower than the cost of govern when the council size has $N$ councilors:

\[
C_G(N+1) \leq C_G(N)
\]

Therefore, to satisfy increasing differences in XXXX, 

\end{proof}

This gives us our first empirical hypothesis to test: if the cost of govern decreases when the legislature increases, then the public goods provision increase when the legislature size increase. But we still need to investigate further how the costs of govern would decrease with legislature size. On the one hand, it seems evident that with the additional legislator, comes also an additional rent and a possible veto player. In the next Lemma we show that the rents always decrease when we increase the number of legislators.

\begin{lemma}
The amount of rents required to win support from a given legislator decreases when the number of legislators increase.
\end{lemma}

\begin{proof}

The proof consists in showing that when we increase the size of the legislature from $N$ to $N+1$, the amount of rents decrease.

\[
\begin{aligned}
b(N+1) - b(N) \ = \ & \dfrac{1}{2(N+1)+1}\left[2\delta R \left(1 + \dfrac{1-\delta}{2(N+1)}\right)-(N+1)(\underline{t} + B) + B\right] \\
& - \dfrac{1}{2N+1}\left[2\delta R \left(1 + \dfrac{1-\delta}{2N}\right)-N(\underline{t} + B) + B\right] \\
= \ & XXXX \\
= \ & XXXX \\
< \ &0
\end{aligned}
\]

\end{proof}

However, the decrease in rents is not sufficient to have lower costs to govern when we increase the number of legislators. The political costs, if high enough, can offset the gains with lower rents. Higher political costs result from having more members of the opposition in the legislature. The next proposition shows the sufficient condition for having a lower cost of govern when we increase the number of legislators.

\begin{proposition}\label{mainprop}

There exists an ex-ante probability of electing a councilor aligned with the government $\overline{\gamma}$ such that, for all $\gamma > \overline{\gamma}$, the cost of govern decreases as $N$ increases, i.e., $C_G(N+1) \leq C_G(N)$.

\end{proposition}

\begin{proof}

Let us rewrite the results as:

\[
XXXX \geq 0
\]

Rearranging the terms:

\[
- \underline{t} - \underline{x}_o^g(N+1) - \dfrac{N}{2}(\underline{x}_o^g(N+1) - \underline{x}_o^g(N-1)) -2p \left[\sum_{\eta_g = 0}^{N/2+1}\mathbb{P}(\eta_g)\eta_g - \sum_{\eta_g = 0}^{N/2}\mathbb{P}(\eta_g)\eta_g\right]
\]

Let $\Delta\eta_n(\gamma) \equiv \left[\sum_{\eta_g = 0}^{N/2+1}\mathbb{P}(\eta_g)\eta_g - \sum_{\eta_g = 0}^{N/2}\mathbb{P}(\eta_g)\eta_g\right]$. There is no closed form solution for this equation. We can approximate the boundaries using @wXXXX but graphically, we can have a good sense of the XXXX:

XXXX GRAPH DELTA GAMMA HERE

And solving for $\Delta\eta_n(\gamma)$ we have that:

\[
XXXX
\]

The natural boundaries for $\Delta\eta_n(\gamma)$ are minus-one and one. Minus-one means that the additional legislature seat went to the opposition while one means that the additional legislature seat went to the government. XXXX 

\end{proof}

Taking stocks, the model provides two testable hypothesis. Let a legislature that previously had $N$ legislators, and now increased the legislature by $N+1$. Then:

\begin{enumerate}
  \item[H1. ] The provision of public goods increase when the costs to govern decrease with legislature size.
  \item[H2. ] The costs to govern decrease with legislature size when the chance of electing a government aligned legislator is sufficiently high.
\end{enumerate}

## Paper Version of the Model

In the main paper, we simplify the analysis with two assumptions regarding the costs of govern. First, we assume a simpler reversal where there is only one reversal stage and after that a reversal policy of zero in all policy components take place. Second, we assume that the cost for govern will depend on the ex-ante expected number of legislator belonging to the government and the opposition. In this case, the rents $b(N)$ become:

\[
XXXX
\]

And Proposition \ref{mainprop} becomes:

\begin{proposition}

XXXX

\[
XXXX
\]

\end{proposition}

\begin{proof}
XXXX
\end{proof}

All the dynamics remain the same but the second hypothesis becomes clearer, with a simplified quantity for the probability of electing a government-aligned politician, $\overline{\gamma}$. The simplified $\overline{\gamma}$ is helpful because it shows that the political concerns require a probability larger than one-half, with a precise amount depending on the XXXX.

\newpage
